---
title: "Project 4"
author: "Yaman Yucel"
date: "2023-04-19"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 

Please answer the following questions assuming the single index model holds. Use your project data in the period 01-Jan-2014 to 01-Jan-2019.

1.  Compute estimates for $\alpha_i$ , $\beta_i$, $\sigma_{e_i} ^2$, i = 1:30 by regressing each stock's return on the S&P500.

```{r}
a_all <- read.csv("stockData.csv", sep=",", header=TRUE) 
a <- a_all[1:60,] 
r <- (a[-1,4:ncol(a)]-a[-nrow(a),4:ncol(a)])/a[-nrow(a),4:ncol(a)] 
r_m <- (a[-1,3]-a[-nrow(a),3])/a[-nrow(a),3]
n_stocks = 30
mean_Rm = mean(r_m)
var_Rm <- var(r_m)
stdev_Rm <- var_Rm^.5

mean_Ri = colMeans(r)

betas = rep(0,n_stocks)
alphas = rep(0,n_stocks)
var_es = rep(0,n_stocks)
var_betas = rep(0,n_stocks)
```

Fit Coefficients

```{r}
for (i in 1:n_stocks){
  fit <- lm(r[,i] ~ r_m)
  betas[i] = fit$coefficients[2]
  alphas[i] = fit$coefficients[1]
  var_es[i] = sum(fit$residuals^2)/ (nrow(r) - 2)
  var_betas[i] = vcov(fit)[2,2]
}

#find beta_i
print("Betas are:")
print(betas)
print("Number of negative betas")
print(sum(which(betas < 0)))

#find alpha_i
print("Alphas are:")
print(alphas)
```
2. Construct the 30x30 variance covariance matrix based on the single index model.

```{r}
#Compute covariance matrix using single index model
covariance_matrix = matrix(0,n_stocks,n_stocks)
for (i in 1:n_stocks)
{
  for(j in 1:n_stocks){
    if(i == j)
    {
      covariance_matrix[i,j] = betas[i] * betas[i] * var_Rm + var_es[i]
    }else{
      covariance_matrix[i,j] = betas[i] * betas[j] * var_Rm
    }
  }
}
print("Single Index Model cov_mat")
print(covariance_matrix)
```
3. Answer the same question as in project 2, part (e) using the new inputs from (1) above. Draw the frontier on
the same plot as in project 2. Now you will have two frontiers, one using the historical variance covariance
matrix (project 2) and one using the variance covariance matrix with inputs from the single index model.

```{r}
#Project 2 part e replicated:
 
n_stocks = 30 
means <- colMeans(r)
covmat <- cov(r)
cormat <- cor(r)
variances <- diag(covmat)
stdev <- diag(covmat)^.5
inv_covmat <- solve(covmat)
ones = rep(1,n_stocks)

A = as.numeric(t(means) %*% inv_covmat %*% ones) 
B = as.numeric(t(means) %*% inv_covmat %*% means) 
C = as.numeric(t(ones)  %*% inv_covmat %*% ones) 
D = B*C - A^2

E <- seq(-0.2,0.2,.001)

sigmas <- sqrt(seq(1/C,0.03,.0001)) 
upper_part <- (A + sqrt(D*(C*sigmas^2 - 1)))*(1/C) 
lower_part <- (A - sqrt(D*(C*sigmas^2 - 1)))*(1/C) 

plot(sigmas, upper_part, lwd=5,type = "l",ylab = 'E', xlab = expression(sigma),ylim = c(-0.15,0.15),main="Risk-Return Plot") 
lines(sigmas,lower_part, lwd=5,type = "l")

#Finding efficient frontier using the single index model:

mean_Ri = colMeans(r)
inv_covmat_single_index = solve(covariance_matrix)
ones = rep(1,n_stocks)
  
A = as.numeric(t(mean_Ri) %*% inv_covmat_single_index %*% ones) 
B = as.numeric(t(mean_Ri) %*% inv_covmat_single_index %*% mean_Ri) 
C = as.numeric(t(ones)  %*% inv_covmat_single_index %*% ones) 
D = B*C - A^2

E <- seq(-0.2,0.2,.001)
sigmas <- sqrt(seq(1/C,0.03,.0001)) 
upper_part <- (A + sqrt(D*(C*sigmas^2 - 1)))*(1/C) 
lower_part <- (A - sqrt(D*(C*sigmas^2 - 1)))*(1/C) 

lines(sigmas, upper_part, lwd=5,type = "l",col = "green") 
lines(sigmas,lower_part, lwd=5,type = "l",col = "green")
legend("topright", 
       legend=c("Historical Var-Cov", "Single index model var-cov"),
       col=c("black", "green"),
       pch = 19,
       fill =c("black", "green"),
       cex=0.8)
```
