---
title: "Project 5"
author: "Yaman Yucel"
date: "2023-04-27"
output:
  pdf_document: default
  html_document: default
---
<br>
Project 5

Before answering questions I have computed betas for periods 01-Jan-2015 to 01-Jan-2020 named as period 1, and 01-Jan-2020 to 31-Mar-2023 named as period 2
<br>
```{r}
a_all <- read.csv("stockData.csv", sep=",", header=TRUE) 

# Use 5 year data to train 
a_period1 <- a_all[1:60,] 
r_period1 <- (a_period1[-1,4:ncol(a_period1)]-a_period1[-nrow(a_period1),4:ncol(a_period1)])/a_period1[-nrow(a_period1),4:ncol(a_period1)] # return of stocks
r_m_period1 <- (a_period1[-1,3]-a_period1[-nrow(a_period1),3])/a_period1[-nrow(a_period1),3] # return of market

#use test data to compute actual predictions
a_period2 <- a_all[61:nrow(a_all),]
r_period2 <- (a_period1[-1,4:ncol(a_period1)]-a_period1[-nrow(a_period1),4:ncol(a_period1)])/a_period1[-nrow(a_period1),4:ncol(a_period1)] # return of stocks
r_m_period2 <- (a_period1[-1,3]-a_period1[-nrow(a_period1),3])/a_period1[-nrow(a_period1),3] # return of market

n_stocks = 30

#Compute period 1 betas
mean_Rm_period1 = mean(r_m_period1)
var_Rm_period1 <- var(r_m_period1)
stdev_Rm_period1 <- var_Rm_period1^.5
mean_Ri_period1 = colMeans(r_period1)

betas_period1 = rep(0,n_stocks)
alphas_period1 = rep(0,n_stocks)
var_es_period1 = rep(0,n_stocks)
var_betas_period1 = rep(0,n_stocks)

for (i in 1:n_stocks){
  fit <- lm(r_period1[,i] ~ r_m_period1)
  betas_period1[i] = fit$coefficients[2]
  alphas_period1[i] = fit$coefficients[1]
  var_es_period1[i] = sum(fit$residuals^2)/ (nrow(r_period1) - 2)
  var_betas_period1[i] = vcov(fit)[2,2]
}
#Compute period 2 betas
mean_Rm_period2 = mean(r_m_period2)
var_Rm_period2 <- var(r_m_period2)
stdev_Rm_period2 <- var_Rm_period2^.5
mean_Ri_period2 = colMeans(r_period2)

betas_period2 = rep(0,n_stocks)
alphas_period2 = rep(0,n_stocks)
var_es_period2 = rep(0,n_stocks)
var_betas_period2 = rep(0,n_stocks)

for (i in 1:n_stocks){
  fit <- lm(r_period2[,i] ~ r_m_period2)
  betas_period2[i] = fit$coefficients[2]
  alphas_period2[i] = fit$coefficients[1]
  var_es_period2[i] = sum(fit$residuals^2)/ (nrow(r_period2) - 2)
  var_betas_period2[i] = vcov(fit)[2,2]
}

```
<br>
Part a: 
Assume the single index model holds. Use only the stocks with positive betas in your data. Choose a value
of Rf and find the optimal portfolio (point of tangency) using the optimization procedure as discussed in
handout #12: http://www.stat.ucla.edu/~nchristo/statistics_c183_c283/statc183c283_tangent.pdf .
<br>
```{r}
#Construct covariance matrix of SIM 
covariance_matrix_period1 = matrix(0,n_stocks,n_stocks)
for (i in 1:n_stocks)
{
  for(j in 1:n_stocks){
    if(i == j)
    {
      covariance_matrix_period1[i,j] = betas_period1[i] * betas_period1[i] * var_Rm_period1 + var_es_period1[i]
    }else{
      covariance_matrix_period1[i,j] = betas_period1[i] * betas_period1[j] * var_Rm_period1
    }
  }
}

inv_covmat_single_index = solve(covariance_matrix_period1)
ones = rep(1,n_stocks)

A = as.numeric(t(mean_Ri_period1) %*% inv_covmat_single_index %*% ones) 
B = as.numeric(t(mean_Ri_period1) %*% inv_covmat_single_index %*% mean_Ri_period1) 
C = as.numeric(t(ones)  %*% inv_covmat_single_index %*% ones) 
D = B*C - A^2

E <- seq(-0.2,0.2,.001)

sigmas <- sqrt(seq(1/C,0.03,.0001)) 
upper_part <- (A + sqrt(D*(C*sigmas^2 - 1)))*(1/C) 
lower_part <- (A - sqrt(D*(C*sigmas^2 - 1)))*(1/C) 

plot(sigmas, upper_part, lwd=5,type = "l",col = "green",xlim = c(0, 0.2), ylim= c(-0.2,0.2)) 
lines(sigmas,lower_part, lwd=5,type = "l",col = "green")

R_f = 0.002
R = mean_Ri_period1 - R_f

Z = inv_covmat_single_index %*% R
x_G = Z / sum(Z)

print("Composition of tangent")
print(x_G)


varg <- t(x_G) %*% covariance_matrix_period1 %*% x_G 
Rg <- t(x_G) %*% mean_Ri_period1 
sigmag <- sqrt(varg)
points(sigmag,Rg, pch=19,lwd=1,col="blue")
abline(a = R_f, b = (Rg - R_f)/sigmag , lwd = 2, col = "red")


legend("topright", 
       legend=c("Single index model var-cov","Tangent when R_f = 0.002", "CAL when R_f = 0.002"),
       col=c("green","blue","red"),
       pch = 19,
       fill =c("green","blue","red"),
       cex=0.8)

```
<br>
Part b: Adjusting the betas:
Adjust the betas using Blume’s and Vasicek’s techniques. For the Blume technique use the two periods: 01-
Jan-2015 to 01-Jan-2020 and 01-Jan-2020 to 31-Mar-2023. For the Vasicek technique use only the period
01-Jan-2014 to 01-Jan-2019.

Note:
For the Blume technique our goal is to adjust the betas in 01-Jan-2020 to 31-Mar-2023 to be better forecasts
for the betas in period 01-Apr-2023 to 01-Apr-2027.

For the Vasicek technique our goal is to adjust the betas in 01-Jan-2015 to 01-Jan-2020 to be better forecasts
for the betas in period 01-Jan-2020 to 31-Mar-2023.
<br>
```{r}
# Blume 
# Use both periods to adjust betas
fit = lm(betas_period2 ~ betas_period1)
beta_adj_blume = fit$coef[1] + fit$coef[2] * betas_period2

print("Adjusted betas using blume technique")
print(beta_adj_blume)

# Vasicek
# In this part period 2 is the forecasted period and period 1 is the historical period
# So we do not use period 2 to adjust betas


vasicek_term_1 = var_betas_period1*mean(betas_period1)/(var(betas_period1)+var_betas_period1)
vasicek_term_2 = var(betas_period1)*betas_period1/(var(betas_period1)+var_betas_period1)
beta_adj_vasicek <- vasicek_term_1 + vasicek_term_2
print("Adjusted betas using vasicek technique")
print(beta_adj_vasicek)
```
<br>
Part c: Compute PRESS only for the Vasicek technique. (You can compute the PRESS only for the Vasicek technique
because you have the actual betas in the period 01-Jan-2020 to 31-Mar-2023.)
<br>
```{r}
PRESS_direct = sum((beta_adj_vasicek - betas_period2)^2)/(n_stocks)

fit <- lm(betas_period2 ~ beta_adj_vasicek)
beta_regress = fit$coefficients[2]
r2 = summary(fit)$r.squared

#Bias Component
term1 = (mean(betas_period2) - mean(beta_adj_vasicek))^2
print("BIAS term:")
print(term1)

Sa2 <- (29/30)*var(betas_period2)
Sp2 <- (29/30)*var(beta_adj_vasicek)

term2 = ((1-as.numeric(beta_regress))^2)*Sp2
print("Inefficiency")
print(term2)
term3 = (1-r2)*Sa2
print("Random error")
print(term3)
PRESS_indirect = term1 + term2 + term3
print("Press with decomposition")
print(PRESS_indirect)
print("Press direct")
print(PRESS_direct)
```